---
title: "Lab 8"
author: "Your Name"
date: "10/11/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Instructions

1.  Before beginning this assignment, rename the file by following these steps:

-   From the Files tab in the bottom right RStudio pane, check the box for `lastname_firstname_lab_08.Rmd`. Then, click Rename.
-   Replace lastname and firstname with your last and first name, followed by `_lab_08.Rmd`. Then click OK.
-   (Alternatively, you can click File --\> Save As. However, this creates a copy of the file, so after saving a new name, you'll need to delete the original file `lastname_firstname_lab_08.Rmd`).

2.  Enter your name for author at the top of this document (where it says `author: "Your Name"`).
3.  If you'd like, use the visual markdown editor by clicking the top-right button of this document (the one that looks sort of like an A).

---

## Comparing base R to tidyverse

So far in this course, you've learned how to manipulate data frames using logical subsetting and other related skills. For this lab, you'll be getting a gentle introduction to data wrangling with dplyr (a tidyverse package). In odd-numbered questions (through 15), you will use base R (what we've already learned). In even-numbered questions (through 16), you will use tidyverse functions.

1. Use the `head` function (base R) to print the first few rows of the `iris` dataset.

```{r}

```

2. Use the `glimpse` function (tidyverse) to print the first few rows of the `iris` dataset.

```{r}

```

3. Extract the rows where the `Species` is `versicolor` (store these results in an object named `dat`). Then, use base R to print the first few rows of `dat`.

```{r}

```

4. Look at slides 1 through 5 on this flipbook example: <https://evamaerey.github.io/data_manipulation/one_stream_wrangle.html>.
      -   On slide 5, notice how they used `%>%` -- that's just the tidyverse way of passing data into a function.
            -   The `%>%` operator is called a *pipe*, and when you use the pipe operator, it's called *piping*.the `filter` function. 
      -   Use the `filter` function to extract the rows of `iris` where the `Species` is `versicolor`. Store it in an object named `dat2`.
            -   Hint: your code should look like this: `dat2 <- iris %>% filter()`. You just have to figure out what to put inside the `filter` function.
      -   On the next line, use tidyverse to print the first few rows of `dat2`.
      
```{r}

```
      
5. In `dat`, `Species` is the 5th column. Use base R to remove the `Species` column from `dat`. Store the results in `dat`, so that `dat` no longer has a `Species` column.
      -   Hint: To remove (instead of subset) a column, use a negative number for the column you are trying to remove.
      -   Then use base R to print the first few rows of `dat`.

```{r}

```

6. Look at slide 6 from the flipbook linked in problem 4. Use the `select` function to remove the `Species` column from `dat2`. Then use tidyverse to print the first few rows of `dat2`.

```{r}

```

7. Use base R (`colnames` function) to rename the 4th column of `dat` (`Petal.Width`) to `pw`. Then use base R to print the first few rows of `dat`.

```{r}

```

8. Look at slides 7 from the flipbook. Use the `rename` function to rename `Petal.Width` to `pw` (store the output in `dat2`). Then use tidyverse to print the first few rows of `dat2`.

```{r}

```

9. In base R, you can add a column to a data frame using the `cbind` function. For example: `dat <- cbind(dat, new_column = x)`. This will create a new column in `dat` called `new_column`, and the column will contain whatever `x` is. (Alternatively, you could do `cbind(dat,x)` and then use `colnames(dat)[5] <- ` to  rename the new column.
      -   Use the `cbind` function to create a new column in `dat` which contains the product of `Sepal.Length` and `Sepal.Width`. The column should be named `pd`.
      -   Use base R to print the first few rows of `dat`.
      
```{r}

```

10. Look at slide 8-10 from the flipbook. Use the `mutate` function to create a new column in `dat2` named pd, which contains the product of `Sepal.Length` and `Sepal.Width`.
      -   Be sure to store the output in `dat2`
      -   Then print the first few rows of `dat2` using tidyverse.

```{r}

```
      
11. Use base R to subset the following columns of `dat`: `Sepal.Length`, `pw`, and `pd`. Store the output in `dat`, and use base R to print the first few rows.
      -   Hint: Use the `c()` function to create either a vector of column numbers OR a character vector with the column names (with quotation marks) to subset these columns.

```{r}

```

12. Look at slide 11 from the flipbook. Use the `select` function to overwrite `dat2` so that it only contains the columns `Sepal.Lenth`, `pw`, and `pd`. Then use tidyverse to print the first few rows of `dat2`.
      -   Note: you can use the column names (without quotation marks) OR the column numbers.

```{r}

```

13. In base R, we can use the `sort` function to sort a vector either alphabetically or in numerical order.
      -   For example, `sort(dat$Sepal.Length)` will sort the Sepal Lengths in the `dat` data frame from lowest to highest.
      -   You can use the `rev` function to reverse a vector. So `rev(sort(dat$Sepal.Length))` would give you these values from highest to lowest.
      -   We can use the `order` function to return the vector positions corresponding to the ordering obtained by sort.
            -   For example, `dat$Sepal.Length[order(dat$Sepal.Length)]` is equivalent to `sort(dat$Sepal.Length)`.
            -   That's not very useful, but it is IS useful in reordering the rows of a data frame -- e.g. `dat[order(x),]`
      -   Use the `order` function to reorder the rows of dat based on `Sepal.Length`, **from highest to lowest**. Store the results in `dat` and print the first few rows using base R.
            -   Hint: for highest to lowest ordering, you can use the `rev` function, or you can put a minus sign in front of `dat$Sepal.Length`. Either one will work.

```{r}

```

14. Look at slide 12 from the flipbook. Observe what the `case_when` function does. It's quite interesting! But we're not going to cover it (at least not today). After studying slide 12, move on to slide 13.
      -   Use the `arrange` function to order `dat2` rows by `Sepal.Length` **from highest to lowest**.
      -   Store the output in `dat2` and use tidyverse to print the first few rows.

```{r}

```

15. Using base R, subset `dat` to include rows 3 through 25. Store the results in `dat`, and use base R to print the first few rows.

```{r}

```

16. Look at slide 15 from the flipbook, and use the `slice` function to subset rows 3 through 25. Store the results in `dat2`, and use tidyverse to print the first few rows.

```{r}

```

17. Starting with the iris dataset, perform all of these changes using a single command. (For stylistic purposes, start a new line every time you use `%>%`)
      -   At every step, run the code to preview the output before you add a new `%>%`
      -   Also, don't store this in an object -- just let the entire object print when you run it
      -   Use the `filter` function to extract the rows of `iris` where the `Species` is `versicolor`.
      -   Use the `select` function to remove the `Species` column
      -   Use the `rename` function to rename `Petal.Width` to `pw`
      -   Use the `mutate` function to create a new column named pd, which contains the product of `Sepal.Length` and `Sepal.Width`
      -   Use the `select` function to to only include the columns `Sepal.Lenth`, `pw`, and `pd`.
      -   Use the `arrange` function to order the rows by `Sepal.Length` from highest to lowest
      -   Use the `slice` function to subset rows 3 through 25

```{r}

```

18. Now for some ggplot2 practice! Using the `iris` dataset and `ggplot2`, make a scatterplot of `Sepal.Length` vs `Sepal.Width`. Color-code the points by `Species`, and make the size of the points propotional to `Petal.Width`.

```{r}

```

19. Make the same plot as you did in 17, but this time:
      -   Don't scale size by `Petal.Width`
      -   Use a different them (anything but `theme_grey()`)
      -   Add a line (one for each species) -- within the function that plots the line, include the arguments `method = "lm"` and `se = FALSE`
      -   Make the points transparent (you decide how transparent you want them to be)
      -   Add a title: "Sepal Length vs Sepal Width"
      -   For the x and y labels, remove the period after Sepal, and also add the units (cm) in parentheses
      
```{r}

```

Save and Knit the document. Be sure the HTML file successfully renders. Then submit BOTH the .Rmd and the HTML file on Webcourses.